SOM result tables

### 2025-01-15 analysis/Supplementary/supplementary_tables.qmd
Supplementary tables improved and organised. WIP render.

---
title: "SOM result tables"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

@@

terrace_RM_WIP <- terrace_RM %>%
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    .default = RMTYPE)) %>% 
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%
  distinct(ID, .keep_all = TRUE)

@@

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset %>% 
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) %>% # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))

@@

shelter_RM_WIP <- shelter_RM %>%
  mutate_at("ID", str_replace, "VB", "") %>% # Remove "VB" from ID
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "TYPE 2",
    RMTYPE == "Chalcedony" ~ "TYPE 1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "TYPE 3",
    RMTYPE == "TYPE 7" ~ "TYPE 4",
    RMTYPE == "TYPE 12" ~ "TYPE 5",
    RMTYPE == "TYPE 11" ~ "TYPE 6",
    RMTYPE == "TYPE 13" ~ "TYPE 7",
    RMTYPE == "TYPE 11F" ~ "TYPE 8",
    RMTYPE == "TYPE 13B" ~ "TYPE 9",
    RMTYPE == "Oolitic" ~ "TYPE 10",
    RMTYPE == "TYPE GB" ~ "TYPE 11",
    RMTYPE == "ND_A" ~ "INDET",
    .default = RMTYPE)) %>%
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%  # Select variables
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets

@@

```{r}
#| label: tbl-bcratio-solutrean
#| tbl-cap: Blank to core ratio by raw material type of the Solutrean assemblage (levels A to C) of the Shelter area.

kable(solutrean_ratiobc)

```{r}
#| label: tbl-bcratio-proto
#| tbl-cap: Blank to core ratio by raw material type of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area.

kable(proto_ratiobc)

```{r}
#| label: tbl-bcratio-gravettian
#| tbl-cap: Blank to core ratio by raw material type of the Gravettian assemblage (levels 6 and 7) of the Terrace area.

kable(gravettian_ratiobc)

@@

```{r}
#| label: tbl-tdratio-solutrean
#| tbl-cap: Tools to debitage ratio by raw material of the Solutrean assemblage (levels A to C) of the Shelter area.

kable(solutrean_ratiotd)

```{r}
#| label: tbl-tdratio-proto
#| tbl-cap: Tools to debitage ratio by raw material of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area.

kable(proto_ratiotd)

```{r}
#| label: tbl-tdratio-gravettian
#| tbl-cap: Tools to debitage ratio by raw material of the Gravettian assemblage (levels 6 and 7) of the Terrace area.

kable(gravettian_ratiotd)

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
solutrean_stats_flakes <- solutrean_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
solutrean_stats_elong <- solutrean_stats %>% 
  filter(CLASS == "ElongatedProd") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
solutrean_stats_cores <- solutrean_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
solutrean_stats_retouched <- solutrean_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
proto_stats_flakes <- proto_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
proto_stats_elong <- proto_stats %>% 
  filter(CLASS == "ElongatedProd") %>% 
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
proto_stats_cores <- proto_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
proto_stats_retouched <- proto_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

@@

>>>>

---
title: "SOM result tables"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: ../paper/references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

```{r}
#| label: library

library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)

@@

terrace_RM_WIP <- terrace_RM %>%
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "T2",
    RMTYPE == "Chalcedony" ~ "T1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "T3",
    RMTYPE == "TYPE 7" ~ "T4",
    RMTYPE == "TYPE 12" ~ "T5",
    RMTYPE == "TYPE 11" ~ "T6",
    RMTYPE == "TYPE 13" ~ "T7",
    RMTYPE == "TYPE 11F" ~ "T8",
    RMTYPE == "TYPE 13B" ~ "T9",
    RMTYPE == "Oolitic" ~ "T10",
    RMTYPE == "TYPE GB" ~ "T11",
    .default = RMTYPE)) %>% 
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%
  distinct(ID, .keep_all = TRUE)

@@

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset |>
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) |> # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))

@@

shelter_RM_WIP <- shelter_RM %>%
  mutate_at("ID", str_replace, "VB", "") %>% # Remove "VB" from ID
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "T2",
    RMTYPE == "Chalcedony" ~ "T1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "T3",
    RMTYPE == "TYPE 7" ~ "T4",
    RMTYPE == "TYPE 12" ~ "T5",
    RMTYPE == "TYPE 11" ~ "T6",
    RMTYPE == "TYPE 13" ~ "T7",
    RMTYPE == "TYPE 11F" ~ "T8",
    RMTYPE == "TYPE 13B" ~ "T9",
    RMTYPE == "Oolitic" ~ "T10",
    RMTYPE == "TYPE GB" ~ "T11",
    RMTYPE == "ND_A" ~ "INDET",
    .default = RMTYPE)) %>%
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%  # Select variables
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets

@@

```{r}
#| label: tbl-bcratio
#| tbl-cap: Blank to core ratio by raw material of the Solutrean (a), Proto-Solutrean (b) and Gravettian (c) assemblages.
#| tbl-subcap: 
#|  - "Blank to core ratio by raw material of the Solutrean assemblage (levels A to C) of the Shelter area."
#|  - "Blank to core ratio by raw material of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area."
#|  - "Blank to core ratio by raw material of the Gravettian assemblage (levels 6 and 7) of the Terrace area."
#| layout-nrow: 4

kable(head(solutrean_ratiobc))
kable(head(proto_ratiobc))
kable(head(gravettian_ratiobc))

@@

```{r}
#| label: tbl-tdratio-solutrean
#| tbl-cap: Tools to debitage ratio by raw material of the Solutrean (a), Proto-Solutrean (b) and Gravettian (c) assemblages.
#| tbl-subcap: 
#|  - "Tools to debitage ratio by raw material of the Solutrean assemblage (levels A to C) of the Shelter area."
#|  - "Tools to debitage ratio by raw material of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area."
#|  - "Tools to debitage ratio by raw material of the Gravettian assemblage (levels 6 and 7) of the Terrace area."
#| layout-nrow: 4

kable(head(solutrean_ratiotd))
kable(head(proto_ratiotd))
kable(head(gravettian_ratiotd))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
solutrean_stats_flakes <- solutrean_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
solutrean_stats_elong <- solutrean_stats %>% 
  filter(CLASS == "ElongatedProd") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
solutrean_stats_cores <- solutrean_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
solutrean_stats_retouch<- solutrean_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

@@

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
proto_stats_flakes <- proto_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
proto_stats_elong <- proto_stats %>% 
  filter(CLASS == "ElongatedProd") %>% 
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
proto_stats_cores <- proto_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
proto_stats_retouch <- proto_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

@@

```{r}
#| label: tbl-solutrean
#| tbl-cap: "Stats for Solutrean occupation."
#| tbl-subcap: 
#|  - "Stats for flakes"
#|  - "Stats for elongated blanks"
#|  - "Stats for cores"
#|  - "Stats for retouched"
#| layout-nrow: 4

kable(head(solutrean_stats_flakes))
kable(head(solutrean_stats_elong))
kable(head(solutrean_stats_cores))
kable(head(solutrean_stats_retouch))

```{r}
#| label: tbl-proto
#| tbl-cap: "Stats for Proto-Solutrean occupation."
#| tbl-subcap: 
#|  - "Stats for flakes"
#|  - "Stats for elongated blanks"
#|  - "Stats for cores"
#|  - "Stats for retouched"
#| layout-nrow: 4

kable(head(proto_stats_flakes))
kable(head(proto_stats_elong))
kable(head(proto_stats_cores))
kable(head(proto_stats_retouch))

```{r}
#| label: tbl-gravettian
#| tbl-cap: "Stats for Gravettian occupation."
#| tbl-subcap: 
#|  - "Stats for flakes"
#|  - "Stats for elongated blanks"
#|  - "Stats for cores"
#|  - "Stats for retouched"
#| layout-nrow: 4

kable(head(gravettian_stats_flakes))
kable(head(gravettian_stats_elong))
kable(head(gravettian_stats_cores))
kable(head(gravettian_stats_retouch))
